An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-30 uai2017/paper.tex
Very minor changes, in appendix mostly

To address this
challenge,~\citep{cutting_mh_2014,icml2014c1_bardenet14} perform
approximate Metropolis-Hasting tests using sequential hypothesis
testing. At each iteration, a subset of data is sampled and used to
test whether to accept $\theta'$ using an approximation to
$\alpha(\theta,\theta')$. If the approximate test does not yield a
decision, the minibatch size is increased and the test repeated. This
process continues until a decision. These methods either invoke the
asymptotic CLT and assume that finite  batch errors are normally
distributed~\citep{cutting_mh_2014} or use a concentration
bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
respectively, as {\sc AustereMH} and {\sc MHSubLhd}. While both show
useful reductions in the number of samples required, they suffer from
two drawbacks: (i) They are approximate, and always yield a decision
with a finite error, (ii) They both require exact, dataset-wide bounds
that depend on $\theta$.
\footnote{We obtained the authors code for both and found that they scanned the entire
dataset at each step to obtain these estimates.} (see
Section~\ref{sec:analysis}). We discuss a
worst-case scenario in Section~\ref{ssec:gaussian_example}.

@@

For the Gaussian mixture experiment, we use the conservative method
from~\citep{cutting_mh_2014}, which avoids the need for recomputing log
likelihoods of each data point each iteration by choosing baseline minibatch
sizes $m$ and per-test thresholds $\epsilon$ beforehand, and then using those
values for the entirety of the trials. We experimented with the following
values:
\begin{itemize}[noitemsep]
\item $\epsilon \in \{0.001, 0.005, 0.01, 0.05, 0.1, 0.2\}$
\item $m \in \{50, 100, 150, 200, 250, 300, 350, 400, 450, 500\}$
\end{itemize}
and chose the $(m,\epsilon)$ pairing which resulted in the lowest expected data
usage given a selected upper bound on the error. Through personal communication
with~\citet{cutting_mh_2014}, we were able to use their same code to compute
expected data usage and errors.

@@

Figure~\ref{fig:logistic_histograms_mnist} shows the histograms for the four
methods on one representative trial of MNIST-13k, which shows similar relative
performance of the four methods as shown in
Figure~\ref{fig:logistic_histograms_mnist8m} (which uses MNIST-100k). In
particular, {\sc MHminibatch} exhibits a shorter-tailed distribution and
consumes nearly an order of magnitude fewer data points compared to {\sc
AustereMH(nc)}, the next-best method; see Table~\ref{tab:logistic} for details.

@@

Table~\ref{tab:logistic_settings} indicates the relevant parameter settings for
the logistic regression experiments. Unless otherwise stated, values apply to
all methods tested. For values from~\citep{cutting_mh_2014}
or~\cite{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
\gamma, p$, and $\delta$) to be consistent.

>>>>

To address this
challenge,~\citep{cutting_mh_2014,icml2014c1_bardenet14} perform
approximate Metropolis-Hasting tests using sequential hypothesis
testing. At each iteration, a subset of data is sampled and used to
test whether to accept $\theta'$ using an approximation to
$\alpha(\theta,\theta')$. If the approximate test does not yield a
decision, the minibatch size is increased and the test repeated. This
process continues until a decision. These methods either invoke the
asymptotic CLT and assume that finite  batch errors are normally
distributed~\citep{cutting_mh_2014} or use a concentration
bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
respectively, as {\sc AustereMH} and {\sc MHSubLhd}. While both show
useful reductions in the number of samples required, they suffer from
two drawbacks: (i) They are approximate, and always yield a decision
with a finite error, (ii) They both require exact, dataset-wide bounds
that depend on $\theta$ (see Section~\ref{sec:analysis}).\footnote{We obtained
the authors code for both and found that they scanned the entire dataset at each
step to obtain these estimates.} We discuss a
worst-case scenario in Section~\ref{ssec:gaussian_example}.

@@

For the Gaussian mixture experiment, we use the conservative method
from~\citep{cutting_mh_2014}, which avoids the need for recomputing log
likelihoods of each data point each iteration by choosing baseline minibatch
sizes $m$ and per-test thresholds $\epsilon$ beforehand, and then using those
values for the entirety of the trials. We experimented with the following
values, which are similar to the values reported in~\citep{cutting_mh_2014}:
\begin{itemize}[noitemsep]
\item $\epsilon \in \{0.001, 0.005, 0.01, 0.05, 0.1, 0.2\}$
\item $m \in \{50, 100, 150, 200, 250, 300, 350, 400, 450, 500\}$
\end{itemize}
and chose the $(m,\epsilon)$ pairing which resulted in the lowest expected data
usage given a selected upper bound on the error. Through personal communication
with~\citet{cutting_mh_2014}, we were able to use their same code to compute
expected data usage and errors.

@@

Figure~\ref{fig:logistic_histograms_mnist} shows the histograms for the four
methods on one representative trial of MNIST-13k, indicating similar relative
performance of the four methods as in
Figure~\ref{fig:logistic_histograms_mnist8m} (which uses MNIST-100k). In
particular, {\sc MHminibatch} exhibits a shorter-tailed distribution and
consumes nearly an order of magnitude fewer data points compared to {\sc
AustereMH(nc)}, the next-best method; see Table~\ref{tab:logistic} for details.

@@

Table~\ref{tab:logistic_settings} indicates the relevant parameter settings for
the logistic regression experiments. Unless otherwise stated, values apply to
all methods tested. For values from~\citep{cutting_mh_2014}
or~\citep{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
\gamma, p$, and $\delta$) to be consistent.
